---
title: "Exploring my music compared to the class corpus"
author: "Desmond Leung"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(flexdashboard)
library(plotly)
library(knitr)
library(kableExtra)
```

### My Corpus

In this project, I analyze my tracks 'desmond-l-1' and 'desmond-l-2' compared to the Computational Musicology 2025 corpus using [Essentia](https://essentia.upf.edu/) features such as *danceability*, *energy*, *valence*, and *instrumentalness.*

I curated my tracks using generative AI tool [Stable Audio](https://stableaudio.com/). The prompts I used were based on two of my favorite artists: [KI/KI](https://open.spotify.com/artist/0UMs6dTf23FC2fHc40fXNS?si=9d1d916c8ac54bb1) and [Marlon Hoffstadt](https://open.spotify.com/artist/0HHa7ZJZxUQlg5l2mB0N0f?si=58729cedd9684ce0). Both DJs are known for their high energetic music, combining acid and trance with techno.

The table below presents the extracted Essentia feature values for my generated tracks.
```{r}
compmus2025 <- read_csv("compmus2025.csv")

mycorpus <- compmus2025 |>
  filter(filename %in% c('desmond-l-1','desmond-l-2'))
  
kable(mycorpus, caption = "Essentia Feature Analysis of My Tracks") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

### The relationship between musical positiveness and emotion

```{r}

mean_arousal <- mean(compmus2025$arousal, na.rm = TRUE)
mean_valence <- mean(compmus2025$valence, na.rm = TRUE)
slope <- mean_valence / mean_arousal  



compmus2025 |>                   # Start with the data
   ggplot(                     # Set up the plot.
    aes(
      x = arousal,
      y = valence,
      colour = danceability,
      label = filename
    )
  ) +
  geom_point() +              # Scatter plot.
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
  # Add diagonal average line
  geom_abline(intercept = 0, slope = slope, linetype = "dashed", color = "red", linewidth = 0.2) +

  geom_text(                  # Add text labels from above.
    x = 5.525241,
    y = 5.837563,
    label = "desmond-l-1",
    size = 3,                 # Override size (not loudness here).
    hjust = "left",           # Align left side of label with the point.
    vjust = "center",         # Align vertical centre of label with the point.
    angle = 30,                # Rotate the text label
  ) +
  geom_text(                  # Add text labels from above.
    x = 6.353199,
    y = 5.886079,
    label = "desmond-l-2",
    hjust = "left",           # Align left side of label with the point.
    vjust = "center",         # Align vertical centre of label with the point.
    angle = 30,                # Rotate the text label
  ) +
  scale_x_continuous(         # Fine-tune the x axis.
    limits = c(3, 7.5),
    breaks = c(3, 4.5, 6, 7.5), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(3, 7.5),
    breaks = c(3, 4.5, 6, 7.5), # Specify grid lines
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # Use the popular viridis colour palette.
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Arousal",
    y = "Valence",
    colour = "Danceability"
  )
ggplotly()
```
***
This graph explores the relationship between arousal and valence, with the danceability represented by the color.

My tracks, **desmond-l-1**, **and desmond-l-2** score high in arousal and valence, meaning they are energetic and also emotionally positive. This aligns perfect with the sound of **KI/KI** and **Marlon Hoffstadt**.

### How rave-ready are my tracks compared to the Class Corpus?

Compared to the class corpus, my tracks exhibit a higher combination of energy, emotional positivity, and danceability. This suggests that they are:

- More suited for raves than the average track in the class corpus.
- More danceable than most class tracks, making them well-suited for a high-energy electronic music setting.



