---
title: "Exploring my music compared to the class corpus"
author: "Desmond Leung"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme:
      bg: "#2D1E40"
      fg: "#F5F5F5"
      primary: "#A3FFBA"
      navbar-bg: "#875F9A"
---

```{r setup, include=FALSE}
library(tidyverse)
library(flexdashboard)
library(plotly)
library(knitr)
library(kableExtra)
source("compmus.R")
```

***
Chromagrams and Cepstrograms  

```{r, out.width="100%", out.height="60%"}

"features/desmond-l-1.json" |>   
  compmus_chroma(norm = "identity") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()            # Change the theme?

"features/desmond-l-1.json" |>   
  compmus_mfccs(norm = "identity") |>                  # Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()                                      # Change the theme?
```

***

Explanation: Chroma & Timbre Analyse

- **Chroma Self-Similarity:** this plot is visualizing the harmonic structure of a musical piece.Patterns in the chroma self similarity can point to repeating chords and music scales. The song chosen is the first track from Berend. This is because my tracks are unfortunatly not yet in the corpus. So I will explain this plot as soon as it is my song. The same goes for the timbre self similarity plot
- **Timbre (MFCC) Self-Similarity:** This represents the timbre of the song. Variations in timbre can be related to instrumentelness and dynamics.

By looking at the plots together we can look at similarities between the plots and maybe see some interesting patterns.

```{r, out.width="100%", out.height="60%"}
"features/desmond-l-1.json" |>                         # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?

"features/desmond-l-1.json" |>                         # Change the track
  compmus_mfccs(norm = "identity") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```

***

Explanation: Chroma & Timbre Analyse

- **Chroma Self-Similarity:** this plot is visualizing the harmonic structure of a musical piece.Patterns in the chroma self similarity can point to repeating chords and music scales. The song chosen is the first track from Berend. This is because my tracks are unfortunatly not yet in the corpus. So I will explain this plot as soon as it is my song. The same goes for the timbre self similarity plot
- **Timbre (MFCC) Self-Similarity:** This represents the timbre of the song. Variations in timbre can be related to instrumentelness and dynamics.

By looking at the plots together we can look at similarities between the plots and maybe see some interesting patterns.
### My Corpus

In this project, I analyze my tracks 'desmond-l-1' and 'desmond-l-2' compared to the Computational Musicology 2025 corpus using [Essentia](https://essentia.upf.edu/) features such as *danceability*, *energy*, *valence*, and *instrumentalness.*

I curated my tracks using generative AI tool [Stable Audio](https://stableaudio.com/). The prompts I used were based on two of my favorite artists: [KI/KI](https://open.spotify.com/artist/0UMs6dTf23FC2fHc40fXNS?si=9d1d916c8ac54bb1) and [Marlon Hoffstadt](https://open.spotify.com/artist/0HHa7ZJZxUQlg5l2mB0N0f?si=58729cedd9684ce0). Both DJs are known for their high energetic music, combining acid and trance with techno.

The table below presents the extracted Essentia feature values for my generated tracks.
```{r}
compmus2025 <- read_csv("compmus2025.csv")

mycorpus <- compmus2025 |>
  filter(filename %in% c('desmond-l-1','desmond-l-2'))
  
kable(mycorpus, caption = "Essentia Feature Analysis of My Tracks") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

### The relationship between musical positiveness and emotion

```{r}

mean_arousal <- mean(compmus2025$arousal, na.rm = TRUE)
mean_valence <- mean(compmus2025$valence, na.rm = TRUE)
slope <- mean_valence / mean_arousal  



compmus2025 |>                   # Start with the data
   ggplot(                     # Set up the plot.
    aes(
      x = arousal,
      y = valence,
      colour = danceability,
      label = filename
    )
  ) +
  geom_point() +              # Scatter plot.
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
  # Add diagonal average line
  geom_abline(intercept = 0, slope = slope, linetype = "dashed", color = "red", linewidth = 0.2) +

  geom_text(                  # Add text labels from above.
    x = 5.525241,
    y = 5.837563,
    label = "desmond-l-1",
    size = 3,                 # Override size (not loudness here).
    hjust = "left",           # Align left side of label with the point.
    vjust = "center",         # Align vertical centre of label with the point.
    angle = 30,                # Rotate the text label
  ) +
  geom_text(                  # Add text labels from above.
    x = 6.353199,
    y = 5.886079,
    label = "desmond-l-2",
    hjust = "left",           # Align left side of label with the point.
    vjust = "center",         # Align vertical centre of label with the point.
    angle = 30,                # Rotate the text label
  ) +
  scale_x_continuous(         # Fine-tune the x axis.
    limits = c(3, 7.5),
    breaks = c(3, 4.5, 6, 7.5), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(3, 7.5),
    breaks = c(3, 4.5, 6, 7.5), # Specify grid lines
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # Use the popular viridis colour palette.
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Arousal",
    y = "Valence",
    colour = "Danceability"
  )
ggplotly()
```
***
This graph explores the relationship between arousal and valence, with the danceability represented by the color.

My tracks, **desmond-l-1**, **and desmond-l-2** score high in arousal and valence, meaning they are energetic and also emotionally positive. This aligns perfect with the sound of **KI/KI** and **Marlon Hoffstadt**.

### How rave-ready are my tracks compared to the Class Corpus?

Compared to the class corpus, my tracks exhibit a higher combination of energy, emotional positivity, and danceability. This suggests that they are:

- More suited for raves than the average track in the class corpus.
- More danceable than most class tracks, making them well-suited for a high-energy electronic music setting.



